Supporting Information

### 2018-03-06 Paper.Rmd
editorial changes

Here we intend to study that if for a given bivariate dataset we are able to plot the data within a unit square box based on their rankings then does that data look like totally independent or do they assume a copula structure? If they do show some dependence then at best by which family of copula we can describe their dependence? To answer these questions, we first choose to pass our bivariate dataset through a p-value based independence-test using *BiCopIndTest* function from *VineCopula* package. For that we set the significance level equal to *0.05*. If independence is rejected then we perform a test based on maximum likelihood estimation which returns us estimated parameters and log-likelihood, AIC, BIC informations for a specific copula indicated by its familycode. We used *BiCopEst* function from *VineCopula* package for families like **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}*. The integer within the curly bracket *`{}`* indicates the familycode of a specific copula. We choose a best-fitted copula based on minimum AIC value and to verify that fitted copula-model we perform a goodness-of-fit test of our given data with that best-fitted copula-family by *BiCopGofTest* function from *VineCopula* package. For the goodness-of-fit test if any of these p-values (based on CvM or KS method) is less than a threshold *(=0.2)* we used *1000* number of bootstraps and *100* otherwise. We also keep the informations like p-values and associated statistics based on Cramer-von Mises test and Kolmogorov-Smirnov test for all next possible fitted copulas in addition to that best-fitted one. We choose a collection of possible good fitted copula by seeting a reasonable bar from the minimum AIC values obtained from bestfitted copula model. Say, if the AIC of any copula is atmost greater than the AIC value of the best-fitted one by a threshold value *(=2 unit)* then we consider that copula to be in the possible good-fit copula list. We can use that same approach for multivariable dataset too. A multivariate dataset when considered as pairwise behaves as a bivariate dataset and we can count how many times a specific copula family wins over others as a bestfit or possible good-fit copula model. If that data has a certain tail dependence then the winners mostly come from either {**C, SG, SJ, SBB8**} or {**G, SC, J, BB8**} category indicating lower and upper tail dependence respectively.

----------------------------------------------------------------------
Shyamolina and Dan to discuss which tail-dependence statistic(s) to use and then we can insert a description here.

>>>>

Here we intend to study that if for a given bivariate dataset we are able to plot the data within a unit square box based on their rankings then does that data look like totally independent or do they assume a copula structure? If they do show some dependence then at best by which family of copula we can describe their dependence? To answer these questions, we first choose to pass our bivariate dataset through a p-value based independence-test using *BiCopIndTest* function from *VineCopula* package. For that we set the significance level equal to *0.05*. If independence is rejected then we perform a test based on maximum likelihood estimation which returns us estimated parameters and log-likelihood, AIC, BIC informations for a specific copula indicated by its familycode. We used *BiCopEst* function from *VineCopula* package for families like **N**ormal/Gaussian*{1}*, **C**layton*{3}*, **G**umbel*{4}*, **F**rank*{5}*, **J**oe*{6}*, **BB1***{7}*, **BB6***{8}*, **BB7***{9}*, **BB8***{10}*, **S**urvival **C**layton*{13}*, **S**urvival **G**umbel*{14}*, **S**urvival **J**oe*{16}*, **S**urvival **BB1***{17}*, **S**urvival **BB6***{18}*, **S**urvival **BB7***{19}* and **S**urvival **BB8***{20}*. The integer within the curly bracket *`{}`* indicates the familycode of a specific copula. We choose a best-fitted copula based on minimum AIC value and to verify that fitted copula-model we perform a goodness-of-fit test of our given data with that best-fitted copula-family by *BiCopGofTest* function from *VineCopula* package. For the goodness-of-fit test if any of these p-values (based on CvM or KS method) is less than a threshold *(=0.2)* we used *1000* number of bootstraps and *100* otherwise. We also keep the informations like p-values and associated statistics based on Cramer-von Mises test and Kolmogorov-Smirnov test for all next possible fitted copulas in addition to that best-fitted one. We choose a collection of possible good fitted copula by seeting a reasonable bar from the minimum AIC values obtained from bestfitted copula model. Say, if the AIC of any copula is atmost greater than the AIC value of the best-fitted one by a threshold value *(=2 unit)* then we consider that copula to be in the possible good-fit copula list. We can use that same approach for multivariable dataset too. A multivariate dataset when considered as pairwise behaves as a bivariate dataset and we can count how many times a specific copula family wins over others as a bestfit or possible good-fit copula model. If that data has a certain tail dependence then the winners mostly come from either {**C, SG, SJ, SBB8**} or {**G, SC, J, BB8**} category indicating lower and upper tail dependence respectively. 
<!--Shya please add some information on model-averaged tail dependence statistics-->
----------------------------------------------------------------------
Shyamolina and Dan to discuss which tail-dependence statistic(s) to use and then we can insert a description here.

### 2018-03-06 SupportingInformation.Rmd
editorial changes

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, fig.cap="plot of the raw soil C and N data.", fig.height=5, fig.width=5}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
fig_nums("soilCNraw",display=F)
plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data",  fig.height=5, fig.width=5}
#convert to a copula and plot
fig_nums("soilCNcop",display=F)
v_CN<-getcopula(d=d,rankon=T,ploton=T) #with real data, rankon should be T
```

@@

### Model selection approach
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
```
The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`.

We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas:
```{r soilCNfit,echo=F}
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
table_nums("soilCNfit",display=F)
knitr::kable(h, caption = "Fitting info for soil C and N dataset. Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe. An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).")
```

`r table_nums("soilCNfit",display="c")` shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

The model-averaged (using AIC) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged (again, AIC) upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.

>>>>

<!--The raw data look like this:-->
```{r load_data_Loecke, echo=F, fig.cap="Plot of the raw soil C and N data.", fig.height=5, fig.width=5}
d<-readRDS("Data/RaCA_soilorganicC_soiltotalN_stocks100cm.RDS")
d<-d[,c("SOCstock100","TSNstock100")]
fig_nums("soilCNraw",display=F)
plot(d[,1],d[,2],type="p",col="red",xlab="Soil C",ylab="Soil N",cex.lab=1)
```

<!--The copula looks like this:-->
```{r rankandplot_Loecke, echo=F, fig.cap="Copula plot of soil C and N data.",  fig.height=5, fig.width=5}
#convert to a copula and plot
fig_nums("soilCNcop",display=F)
v_CN<-getcopula(d=d,rankon=T,ploton=T) 
```

@@

### Model selection approach
<!--This does the computations for this section-->
```{r bivmodselect_Loecke, echo=F, cache=T, cache.extra=list(seed,v_CN,mtime("OurBiCopSelect.R"))}
set.seed(seed)
families_Loecke<-c(1,3:10,13:14,16:20) 
BivMS_res_Loecke<-OurBiCopSelect(u1=v_CN[,1],u2=v_CN[,2],families=families_Loecke,level=0.05,AICBIC="AIC",numBSsmall=100,pthresh=0.2,numBSlarge=1000,gofnormal=F,status=F)
```
The test of indepedence gave $p=$ `r BivMS_res_Loecke$IndepTestRes`. We fitted several copulas
and got AIC values and upper- and lower-tail dependency estimates based on the fitted copulas (`r table_nums("soilCNfit",display="c")`). The table shows that several copulas are much better supported (higher AIC and BIC weights) than the Gaussian copula, answering the first part of question 1 for these data.

<!--Display model selection results in a table-->
```{r soilCNfit,echo=F}
h<-BivMS_res_Loecke$InfCritRes[,c("copname","AICw","BICw","LTdep","UTdep")]
table_nums("soilCNfit",display=F)
knitr::kable(h, caption = "Fitting info for soil C and N dataset. Here and henceforth copula names are abbreviated as: N=normal (Gaussian); C=Clayton; G=Gumbel; F=Frank; J=Joe; SC=Survival Clayton; SG=Survival Gumbel; SJ=Survival Joe. An S in front of one of the copulas BB1, BB6, BB7, BB8 indicates the survival version of that copula (rotation by 180 degrees).")
```

The model-averaged (AIC weights used for model averaging) lower-tail dependence statistic was `r BivMS_res_Loecke$relLTdep_AICw` and the model-averaged upper-tail dependence statistic was `r BivMS_res_Loecke$relUTdep_AICw`. These values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper), `r BivMS_res_Loecke$relLTdep_AICw-BivMS_res_Loecke$relUTdep_AICw`, was different from 0, answering the third part of question 1.
